---
title: "first analysis"
author: "Jan Hengeveld"
date: "2023-03-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, setup for first analysis}

library(tidyverse)
library(plotly)
library(spotifyr)
library(compmus)
```
```{r, selecting data from spotify and combining for analysis}

dogs <- get_playlist_audio_features("", "5hQo2asoxqQrnJFeufycj1")
humans <- get_playlist_audio_features("", "3B0FtfxNiFOo82o8lmJcIp")
dogsscience <- get_playlist_audio_features("", "0km3mDUsP3LYDS1BZfqsY5")
humansscience <- get_playlist_audio_features("", "1t06IDDtn5eYo4Ow7Fwmcb")
combine4lists <- bind_rows(
humans |> mutate(category = "Humans"),
dogs |> mutate(category = "Dogs"),
dogsscience |> mutate(category = "Dogsscience"),
humansscience |> mutate(category = "Humansscience")
)
``

### A global analysis of 4 relaxation playlists: humans, dogs, 'humansscience' and 'dogsscience'

**I started the analysis with a comparison on single elements between the 4 groups**
  

(N.B. at the moment this is still a garbage tab which requires clean-up)

<https://open.spotify.com/playlist/48fYbZUMrk5mJwhfj3xAi4?si=e4adb9f034634348>

------------------------------------------------------------------------

```{r, 1 single variable compared}
combine4lists |>
  ggplot(aes(x = energy)) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(~category)
combine4lists |>
  ggplot(aes(x = category, y = energy)) +
  geom_boxplot()
combine4lists |>
  ggplot(aes(x = category, y = energy)) +
  geom_violin()
```

```{r, 1 single variable compared; tempo}
combine4lists |>
  ggplot(aes(x = tempo)) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(~category)
combine4lists |>
  ggplot(aes(x = category, y = tempo)) +
  geom_boxplot()
combine4lists |>
  ggplot(aes(x = category, y = tempo)) +
  geom_violin()
```

```{r, 1 single variable compared; speechiness}
combine4lists |>
  ggplot(aes(x = tempo)) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(~category)
combine4lists |>
  ggplot(aes(x = category, y = speechiness)) +
  geom_boxplot()
combine4lists |>
  ggplot(aes(x = category, y = speechiness)) +
  geom_violin()
```

```{r, 1 single variable compared; valence}
combine4lists |>
  ggplot(aes(x = valence)) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(~category)
combine4lists |>
  ggplot(aes(x = category, y = valence)) +
  geom_boxplot()
combine4lists |>
  ggplot(aes(x = category, y = valence)) +
  geom_violin()
```

```{r, 1 single variable compared; mode}
combine4lists |>
  ggplot(aes(x = tempo)) +
  geom_histogram(binwidth = 0.1) +
  facet_wrap(~category)
combine4lists |>
  ggplot(aes(x = category, y = mode)) +
  geom_boxplot()
combine4lists |>
  ggplot(aes(x = category, y = mode)) +
  geom_violin()
```

------------------------------------------------------------------------

```{r, scatterplots}

dogs |> ggplot(aes(x = tempo, y = energy)) + geom_point() + geom_smooth()
## `geom_smooth()` using method = 'loess' and formula = 'y ~ x'


```

```{r, code from ashley}
dogs |>                    # Start with awards.
  mutate(
    mode = ifelse(mode == 0, "Minor", "Major")
  ) |>
  ggplot(                     # Set up the plot.
    aes(
      x = loudness,
      y = energy,
      size = valence,
      colour = mode
    )
  ) +
  geom_point() +              # Scatter plot.
  geom_rug(linewidth = 0.1) + # Add 'fringes' to show data distribution.
  geom_text(                  # Add text labels from above.
    aes(
      x = valence,
      y = energy,
      label = label
    ),
    data = 
      tibble(
        label = c("Outlier", "ENERGY"),
        category = c("Edisons", "Grammys"),
        valence = c(0.090, 0.123),
        energy = c(0.101, 0.967)
      ),
    colour = "black",         # Override colour (not mode here).
    size = 3,                 # Override size (not loudness here).
    hjust = "left",           # Align left side of label with the point.
    vjust = "bottom",         # Align bottom of label with the point.
    nudge_x = -0.05,          # Nudge the label slightly left.
    nudge_y = 0.02            # Nudge the label slightly up.
  ) +
  facet_wrap(~ category) +    # Separate charts per playlist.
  scale_x_continuous(         # Fine-tune the x axis.
    limits = c(0, 1),
    breaks = c(0, 0.50, 1),   # Use grid-lines for quadrants only.
    minor_breaks = NULL       # Remove 'minor' grid-lines.
  ) +
  scale_y_continuous(         # Fine-tune the y axis in the same way.
    limits = c(0, 1),
    breaks = c(0, 0.50, 1),
    minor_breaks = NULL
  ) +
  scale_colour_brewer(        # Use the Color Brewer to choose a palette.
    type = "qual",            # Qualitative set.
    palette = "Paired"        # Name of the palette is 'Paired'.
  ) +
  scale_size_continuous(      # Fine-tune the sizes of each point.
    trans = "exp",            # Use an exp transformation to emphasise loud.
    guide = "none"            # Remove the legend for size.
  ) +
  theme_light() +             # Use a simpler theme.
  labs(                       # Make the titles nice.
    x = "loudness",
    y = "Energy",
    colour = "Mode"
  )

```

```{r, code from ashley2}
humans |>                    # Start with awards.
  mutate(
    mode = ifelse(mode == 0, "Minor", "Major")
  ) |>
  ggplot(                     # Set up the plot.
    aes(
      x = valence,
      y = energy,
      size = loudness,
      colour = mode
    )
  ) +
  geom_point() +              # Scatter plot.
  geom_rug(linewidth = 0.1) + # Add 'fringes' to show data distribution.
  geom_text(                  # Add text labels from above.
    aes(
      x = valence,
      y = energy,
      label = label
    ),
    data = 
      tibble(
        label = c("OUtlier", "ENERGY"),
        category = c("Edisons", "Grammys"),
        valence = c(0.090, 0.123),
        energy = c(0.101, 0.967)
      ),
    colour = "black",         # Override colour (not mode here).
    size = 3,                 # Override size (not loudness here).
    hjust = "left",           # Align left side of label with the point.
    vjust = "bottom",         # Align bottom of label with the point.
    nudge_x = -0.05,          # Nudge the label slightly left.
    nudge_y = 0.02            # Nudge the label slightly up.
  ) +
  facet_wrap(~ category) +    # Separate charts per playlist.
  scale_x_continuous(         # Fine-tune the x axis.
    limits = c(0, 1),
    breaks = c(0, 0.50, 1),   # Use grid-lines for quadrants only.
    minor_breaks = NULL       # Remove 'minor' grid-lines.
  ) +
  scale_y_continuous(         # Fine-tune the y axis in the same way.
    limits = c(0, 1),
    breaks = c(0, 0.50, 1),
    minor_breaks = NULL
  ) +
  scale_colour_brewer(        # Use the Color Brewer to choose a palette.
    type = "qual",            # Qualitative set.
    palette = "Paired"        # Name of the palette is 'Paired'.
  ) +
  scale_size_continuous(      # Fine-tune the sizes of each point.
    trans = "exp",            # Use an exp transformation to emphasise loud.
    guide = "none"            # Remove the legend for size.
  ) +
  theme_light() +             # Use a simpler theme.
  labs(                       # Make the titles nice.
    x = "loudness",
    y = "Energy",
    colour = "Mode"
  )

```





